%% ===========================================================================
%%  Chapter 7 — ब्रम्ह Live-History Tracker
%% ===========================================================================
\section{ब्रम्ह Live-History Tracker}\label{ch:tracker}

\texttt{assembly/sakum\_tracker.s} (x86-64) and
\texttt{sakum\_tracker\_arm64*.s} (AArch64) are the ब्रम्ह
(\emph{Brahma}) live self-update viewers, written in raw machine code with
no host language.  They read \texttt{query\_logs/fetch\_live.jsonl} (the
live ledger) and print \texttt{स्रोत → भाषा → गंतव्य} (source $\to$
language $\to$ destination) plus a pulse clock.  All back ends share
identical behaviour.

\subsection{Modes}

\begin{itemize}
  \item \texttt{/tmp/tracker} --- print the current feed once.
  \item \texttt{/tmp/tracker --live} --- tail the feed, re-reading every 3 s
    (\texttt{\_sleep}).
  \item \texttt{/tmp/tracker <path>} --- read a custom feed path.
\end{itemize}

\subsection{Read + line walk (x86-64)}

\begin{lstlisting}[language=SakumAsm,caption={Feed read + line split (sakum\_tracker.s:182).}]
    lea  rdi, [rip + gbuf]
    mov  rsi, 1
    mov  rdx, BUFSZ-1
    mov  rcx, r12
    call _fread
    ...
.walk:
    movzx eax, byte ptr [r15]
    test al, al
    jz   .lastline
    cmp  al, 10
    jne  .wadv
    mov  byte ptr [r15], 0
    mov  rdi, r14
    call print_line
\end{lstlisting}

\subsection{Cross-ISA ports}

\begin{itemize}
  \item \textbf{AArch64 NEON}: vectorises the newline scan with
    \texttt{ld1} / \texttt{dup} / \texttt{cmeq} (16 lanes) / \texttt{cnt};
    output is byte-identical to the scalar arm64 tracker.  This is what
    \texttt{tools/sakum\_tracker.sh} builds on Apple Silicon.
  \item \textbf{ARMv7-A} (\texttt{\_sys} variant): libc-free, \texttt{svc
    \#0} (openat=56, read=63, write=64, close=57, exit=93).
  \item \textbf{RV64GC} (\texttt{\_sys} / \texttt{\_rvv} variants):
    libc-free \texttt{ecall}; the RVV variant uses \texttt{vsetvli} /
    \texttt{vle8.v} / \texttt{vmseq.vx} / \texttt{vfirst.m}.
\end{itemize}

\noindent These ELFs link and are valid, but execute only under user-mode
QEMU or real hardware (absent on this Mac).
